\section*{Authors}

\noindent The SPECFEM3D package was first developed by Dimitri
Komatitsch and Jean-Pierre Vilotte at Institut de Physique du Globe
(IPGP) in Paris, France from 1995 to 1997 and then by Dimitri Komatitsch
and Jeroen Tromp at Harvard University and Caltech, USA, starting in 1998.
The story started on April 4, 1995, when Prof. Yvon Maday from CNRS and University of Paris, France, gave a lecture to
Dimitri Komatitsch and Jean-Pierre Vilotte at IPG about the nice properties of the Legendre spectral-element method with diagonal mass matrix that he had used for
other equations. We are deeply indebted and thankful to him for that.
That followed a visit by Dimitri Komatitsch to OGS (Istituto Nazionale di Oceanografia e di Geofisica Sperimentale) in Trieste, Italy, in February 1995
to meet with G\'eza Seriani and Enrico Priolo, who introduced him to their 2D Chebyshev version of the spectral-element method with a non-diagonal mass matrix.
We are deeply indebted and thankful to them for that.\newline

%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
%%%%%%%% please do NOT merge the lines of the paragraph below, it makes it difficult to insert new names in alphabetical order; leave a single name per line
Since then it has been developed and maintained by a development team: in alphabetical order,
Michael Afanasiev,
Dmitry Alexeev,
Jean-Paul (Pablo) Ampuero,
Kazuto Ando,
\'Etienne Bachmann
Kangchen Bai,
Wolfgang Bangerth,
Piero Basini,
Stephen Beller,
C\'eline Blitz,
Alexis Bottero,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Caio Ciardelli,
Paul Cristini,
Congyue Cui,
Cl\'ement Durochat,
Armando Espindola-Carmona,
Percy Galvez,
Rene Gassm\"oller,
Hom Nath Gharti,
Dominik G\"oddeke,
Sunny Gogar,
Leopold Grinberg,
Vala Hj\"orleifsd\'ottir,
Quancheng Huang,
Elodie Kendall,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Piero Lanucara,
Nicolas Le Goff,
Pieyre Le Loher,
Matthieu Lefebvre,
Wenjie Lei,
Qinya Liu,
Youshan Liu,
David Luet,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Ren\'e Matzen,
Amanda McPherson,
Dennis McRitchie,
Yellapragada Meenakshi,
Jean-Fran\c{c}ois M\'ehaut,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Takayuki Miyoshi,
Vadim Monteiller,
Surendra Nadh Somala,
Tarje Nissen-Meyer,
Andy Nowacki,
Ridvan Orsvuran,
Laura Parisi,
Daniel Peter,
Norbert Podhorszki,
Kevin Pouget,
S.T.G. Raghukanth,
Max Rietmann,
Vittorio Ruggiero,
Elliott Sales de Andrade,
Brian Savage,
Malte Schirwon,
Bernhard Schuberth,
Anne Sieminski,
James Smith,
Leif Strand,
Kai Tao,
Carl Tape,
Julien Thurin,
Jeroen Tromp,
Seiji Tsuboi,
Eduardo Valero Cano,
Brice Videau,
Jean-Pierre Vilotte,
Zhinan Xie,
Chang-Hua Zhang,
Hejun Zhu.\newline


The cover graphic of the manual was created by Santiago Lombeyda from
Caltech's Center for Advanced Computing Research (CACR), USA, with free satellite clipart pictures
from \url{http://www.4vector.com} and \url{http://www.clker.com} added to it.\newline

The code is released open-source under the GNU version 3 license, see the license at the end of this manual.\newline

